---
title: "Analysis of Energy Demand in Victoria"
output: html_notebook
author: "Kristian James Guinto (s3826723)"
---

## Library Imports
```{r}
library(tidyr)
library(dplyr)
library(lubridate)
```

## Import and Merge Energy Data sets
```{r}
# intialise dataframe with just column headers
demand <- data.frame(REGION=NA,
                     SETTLEMENTDATE=NA,
                     TOTALDEMAND=NA,
                     RRP=NA,
                     PERIODTYPE=NA)

# load data from 2014 to 2018
dir <- getwd()
for(state in c("/VIC/", "/NSW/")){
  for(y in c("2014","2015","2016","2017","2018")){
    for(m in seq(1,12)){ # load data for all months
      if(m < 10) { mon <- paste0("0",as.character(m)) }
      else { (mon <- as.character(m)) }
      
      fName = paste0(dir,state,y,"_",mon,".csv")
      tempDf <- read_csv(fName)
      demand <-  rbind(demand, tempDf)
    }
  }
}
demand %>% head()
demand %>% dim()
```

```{r}
dir <- getwd()
path <- paste0(dir,"/VIC/2020_08.csv")
trial <- read_csv(path)
trial %>% head()
```

## Import weather data sets
```{r}
path <- paste0(getwd(),"/WeatherData/")
# load the 4 data sets for weather
maxTemp <- read_csv(paste0(path,"max_temp_melbourne.csv"))
maxTemp %>% head()
```

## Import holiday data sets
source: https://data.gov.au/dataset/ds-dga-b1bc6077-dadd-4f61-9f8c-002ab2cdff10/details
```{r}

```


```{r}
mobility_AU[-1] %>%
  gather(key = "key", value = "value", -Date) %>%
  ggplot(aes(x=Date, y=value)) +
  geom_line() + facet_grid(rows = vars(key))
  
```

```{r}
# helper function replace missing with mean of previous and next non null values
fillMeanAdj <- function(d){
  for(i in seq(from=1,to=length(d),by=1)){
    if(is.na(d[i])){
      x=1
      prevVal <- d[i-x]
      while(is.na(prevVal)){
        prevVal <- d[i-x]
        x = x + 1
      }
      x=1
      nextVal <- d[i+x]
      while(is.na(nextVal)){
        nextVal <- d[i+x]
        x = x + 1
      }
      
      d[i] = (nextVal + prevVal)/2
    }
  }
  return(d);
}
```

### Non-decreasing value
A check for non-decreasing value will also be conducted since the numbers on the data set represent cumulative numbers. This will be done by first taking the lag-1 difference for all variables then checking if these differences are negative. For a non-decreasing sequence, all lag-1 difference must be positive. The results will be summarised by counting the number of negative lag-1 difference per column
```{r}
# Detection of non-decreasing sequence
# helper function to create lag-1 difference. Na's are filled with 0's
daily_1 <- function(x){
  if(is.numeric(x)){
    d <- x - lag(x) # subtracts the data with a lagged 1 copy of the data
    d[is.na(d)] <- 0 # Na is expected on first row, fill this with 0
    return (d)
  }
}
# helper function to summarise results by getting the sum of negative boolean mask
sum.negative <- function(x){
  return (sum(x < 0))
}

is.negative <- function(x){
  return(x < 0)
}

lag1_diff <- covid_AU %>% 
  select(NSW:Total_Deaths) %>% # select numerical variables
  lapply(daily_1) # compute for lag-1 difference

sapply(lag1_diff, sum.negative) # get count of negative numbers per column

```
```{r}
diff(covid_AU$NSW)
```


```{r}
covid_AU$NSW[diff(covid_AU$NSW) < 0] <- NA
```

```{r}
covid_AU <- covid_AU %>% fill(NSW, .direction = "up")
```


```{r}
neg.mask <- sapply(lag1_diff, is.negative)
covid_AU[-1][neg.mask] <- NA
```

```{r}
colSums(is.na(covid_AU))
```

***
To correct instances of decreasing cumulative numbers, the absolute negative lag-1 differences will be added to the respective variables. 

```{r}
# Correction
covid_AU <- covid_AU %>% 
  fill(NSW:Total_Deaths, .direction = "up")
  
```

```{r}

```



